---
title: Welcome Gallery
author: Brett Johnson
date: '2020-05-09'
slug: welcome
categories: []
tags: []
type: ''
subtitle: ''
image: ''
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
```

Welcome, friends, to a data blog by the Hakai Institute.

Explore environmental data with scientists as they seek to uncover the secrets of our complex coastal ecosystems.

Below are some examples of data visualizations you can work on in our [GitHub Repository](http://github.com/hakaiinstitute/open-data-mgmt/Blog)

I really hope you like what you find, and you consider making your work public so that you and others can benefit from collaboration. 

# Hierarchical

```{r}
library(networkD3)
hc <- hclust(dist(USArrests), "ave")

dendroNetwork(hc, height = 600)
```


# Network

```{r, network d3}
library(visNetwork)
data(MisLinks, MisNodes)
plot <- forceNetwork(Links = MisLinks, Nodes = MisNodes, Source = "source",
             Target = "target", Value = "value", NodeID = "name",
             Group = "group", opacity = 0.4)

plot

```

## Sankey

```{r, sankey}
URL <- paste0(
        "https://cdn.rawgit.com/christophergandrud/networkD3/",
        "master/JSONdata/energy.json")
Energy <- jsonlite::fromJSON(URL)
# Plot
sankeyNetwork(Links = Energy$links, Nodes = Energy$nodes, Source = "source",
             Target = "target", Value = "value", NodeID = "name",
             units = "TWh", fontSize = 12, nodeWidth = 30)
```


## Les Miserables

Ever wonder which characters interact with which in the class Les Miserables?

Me neither.

But here it is!


I'd rather see connection among zoop taxa or microbes

```{r, network miserab}
nodes <- jsonlite::fromJSON("https://raw.githubusercontent.com/datastorm-open/datastorm-open.github.io/master/visNetwork/data/nodes_miserables.json")

edges <- jsonlite::fromJSON("https://raw.githubusercontent.com/datastorm-open/datastorm-open.github.io/master/visNetwork/data/edges_miserables.json")


visNetwork(nodes, edges, height = "700px", width = "100%") %>%
  visOptions(selectedBy = "group", 
             highlightNearest = TRUE, 
             nodesIdSelection = TRUE) %>%
  visPhysics(stabilization = FALSE)
```


# Genomics

```{r, eval = FALSE}
library(BioCircos)
library(RColorBrewer)
library(grDevices)

# Define a custom genome
genomeChr = LETTERS
lengthChr = 5*1:length(genomeChr)
names(lengthChr) <- genomeChr

tracks = BioCircosTracklist()
# Add one track for each chromosome
for (i in 1:length(genomeChr)){
  # Define histogram/bars to be displayed
  nbBars = lengthChr[i] - 1
  barValues = sapply(1:nbBars, function(x) 10 + nbBars%%x)
  barColor = colorRampPalette(brewer.pal(8, "YlOrBr"))(length(genomeChr))[i]
  # Add a track with bars on the i-th chromosome
  tracks = tracks + BioCircosBarTrack(paste0("bars", i), chromosome = genomeChr[i], 
    starts = (1:nbBars) - 1, ends = 1:nbBars, values = barValues, color = barColor, 
    range = c(5,75))
}

# Add background
tracks = tracks + BioCircosBackgroundTrack("bars_background", colors = "#2222EE")

BioCircos(tracks, genomeFillColor = "YlOrBr", genome = as.list(lengthChr), 
  genomeTicksDisplay = F, genomeLabelDy = 0)

```


# Time Series

```{r, timeseries}
library(dygraphs)
dygraph(nhtemp, main = "New Haven Temperatures") %>% 
  dyRangeSelector(dateWindow = c("1920-01-01", "1960-01-01"))
```

# Geospatial

