---
title: Welcome
author: Brett Johnson
date: '2020-05-09'
slug: welcome
categories: []
tags: []
type: ''
subtitle: ''
image: ''
---


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
```

Welcome, friends, to a data blog by the Hakai Institute.

Explore environmental data with scientists as they seek to uncover the secrets of our complex coastal ecosystems.

Below are some examples of data visualizations you can work on in our [GitHub Repository](http://github.com/hakaiinstitute/open-data-mgmt/Blog)

I really hope you like what you find, and you consider making your work public so that you and others can benefit from collaboration. 

```{r}
library(networkD3)
data(MisLinks, MisNodes)
plot <- forceNetwork(Links = MisLinks, Nodes = MisNodes, Source = "source",
             Target = "target", Value = "value", NodeID = "name",
             Group = "group", opacity = 0.4)

plot

```

```{r}
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/hobbs-pearson-trials.csv")

fig <- plot_ly(
    type = 'scatterpolargl',
    mode = 'markers'
  )

j = 1
k = 2
for (i in 1:(length(df)/2)){
 fig <- add_trace(
    fig,
    r = df[,j],
    theta = df[,k],
    name = paste('Trial ', i),
    marker = list(
      size = 15,
      line = list(
        color = '#FFF'
      ),
      opacity = 0.7
    )
  )
  j <- j + 2
  k <- k + 2
}

fig <- layout(
  fig,
  title = "Hobbs-Pearson Trials",
  showlegend = F,
  paper_bgcolor = "rgb(223, 223, 223)",
  polar = list(
    bgcolor = "rgb(223, 223, 223)",
    angularaxis = list(
      tickwidth = 2,
      linewidth = 3,
      layer = 'below traces'
    ),
    radialaxis = list(
      side = 'counterclockwise',
      showline = T,
      linewidth = 2,
      tickwidth = 2,
      gridcolor = '#FFF',
      gridwidth = 2
    )
  )
)

fig
```

```{r}
library(plotly)

fig <- plot_ly(
  type = 'contour',
  z = matrix(c(10, 10.625, 12.5, 15.625, 20, 5.625, 6.25, 8.125, 
               11.25, 15.625, 2.5, 3.125, 5, 8.125, 12.5, 0.625, 
               1.25, 3.125, 6.25, 10.625, 0, 0.625, 2.5, 5.625, 
               10), nrow=5, ncol=5),
  colorscale = 'Jet',
  autocontour = F,
  contours = list(
    start = 0,
    end = 8,
    size = 2
  )
)

fig
```

```{r}
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/bcdunbar/datasets/master/parcoords_data.csv")

fig <- df %>%
  plot_ly(width = 1000, height = 600) 
fig <- fig %>% add_trace(type = 'parcoords',
          line = list(color = ~colorVal,
                      colorscale = 'Jet',
                      showscale = TRUE,
                      reversescale = TRUE,
                      cmin = -4000,
                      cmax = -100),
          dimensions = list(
            list(range = c(~min(blockHeight),~max(blockHeight)),
                 constraintrange = c(100000,150000),
                 label = 'Block Height', values = ~blockHeight),
            list(range = c(~min(blockWidth),~max(blockWidth)),
                 label = 'Block Width', values = ~blockWidth),
            list(tickvals = c(0,0.5,1,2,3),
                 ticktext = c('A','AB','B','Y','Z'),
                 label = 'Cyclinder Material', values = ~cycMaterial),
            list(range = c(-1,4),
                 tickvals = c(0,1,2,3),
                 label = 'Block Material', values = ~blockMaterial),
            list(range = c(~min(totalWeight),~max(totalWeight)),
                 visible = TRUE,
                 label = 'Total Weight', values = ~totalWeight),
            list(range = c(~min(assemblyPW),~max(assemblyPW)),
                 label = 'Assembly Penalty Weight', values = ~assemblyPW),
            list(range = c(~min(HstW),~max(HstW)),
                 label = 'Height st Width', values = ~HstW),
            list(range = c(~min(minHW),~max(minHW)),
                 label = 'Min Height Width', values = ~minHW),
            list(range = c(~min(minWD),~max(minWD)),
                 label = 'Min Width Diameter', values = ~minWD),
            list(range = c(~min(rfBlock),~max(rfBlock)),
                 label = 'RF Block', values = ~rfBlock)
            )
          )


fig
```


```{r}
library(dygraphs)
dygraph(nhtemp, main = "New Haven Temperatures") %>% 
  dyRangeSelector(dateWindow = c("1920-01-01", "1960-01-01"))
```

```{r}
library(d3heatmap)
d3heatmap(mtcars, scale="column", colors="Blues")
```

```{r}
URL <- paste0(
        "https://cdn.rawgit.com/christophergandrud/networkD3/",
        "master/JSONdata/energy.json")
Energy <- jsonlite::fromJSON(URL)
# Plot
sankeyNetwork(Links = Energy$links, Nodes = Energy$nodes, Source = "source",
             Target = "target", Value = "value", NodeID = "name",
             units = "TWh", fontSize = 12, nodeWidth = 30)
```

```{r, eval = FALSE}
URL <- paste0(
        "https://cdn.rawgit.com/christophergandrud/networkD3/",
        "master/JSONdata//flare.json")

## Convert to list format
Flare <- jsonlite::fromJSON(URL, simplifyDataFrame = FALSE)

# Use subset of data for more readable diagram
Flare$children = Flare$children[1:3]

radialNetwork(List = Flare, fontSize = 10, opacity = 0.9)
```

```{r}
hc <- hclust(dist(USArrests), "ave")

dendroNetwork(hc, height = 600)
```

