---
title: "Data Management Workshop for the Pelagic Ecosystems Lab and Hakai Affiliates at UBC"
author: "Brett Johnson & Julian Gan"
date: "31/01/2020"
output: bookdown::pdf_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE,echo = TRUE)
library(tidyverse)
library(lubridate)
library(here)
```

# Intermediate Topics in R

## Before we start

https://datacarpentry.org/R-ecology-lesson/00-before-we-start.html#getting_set_up

## Review Introduction to R

### Learning Objectives
* Define the following terms as they relate to R: object, assign, call, function, arguments, options.
* Assign values to objects in R.
* Learn how to name objects
* Use comments to inform script.
* Solve simple arithmetic operations in R.
* Call functions and use arguments to change their default options.
* Inspect the content of vectors and manipulate their content.
* Subset and extract values from vectors.
* Analyze vectors with missing data.


## R Markdown


------------

> ### Learning Objectives
> * Understand what RMarkdown is and why you should use it
> * Learn how to construct an RMarkdown file
> * Export an RMarkdown file into many file formats
------------

### What is R Markdown?

R Markdown allows you to create documents that serve as a neat record of your analysis. In the world of reproducible research, we want other researchers to easily understand what we did in our analysis, otherwise nobody can be certain that you analysed your data properly. You might choose to create an R Markdown document as an appendix to a paper or project assignment that you are doing, upload it to an online repository such as Github, or simply to keep as a personal record so you can quickly look back at your code and see what you did. R Markdown presents your code alongside its output (graphs, tables, etc.) with conventional text to explain it, a bit like a notebook.

#### R Markdown Document Types: 
* Analysis report templates (html, .pdf, or .doc outputs);
* A Manuscript;
* A Book;
* A Dissertation;
* A Research Compendium;
* A Slideshow;
* An interactive dashboard;
* An R Package
* A website

[RMarkdown definitive guide](https://bookdown.org/yihui/rmarkdown/)

[R Markdown Cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/)

Use the [redoc package](https://noamross.github.io/redoc/) for importing tracked changes 

### Create an R Markdown file

To create a new R Markdown file (`.Rmd`), select _File -> New File -> R Markdown_ in RStudio, then choose the file type you want to create. For now we will focus on a HTML document, which can be easily converted to other file types later.

### The different parts of an R Markdown file

#### The YAML Header

At the top of any R Markdown script is a **YAML header** section enclosed by `` --- ``. By default this includes a title, author, date and the file type you want to output to. Many other options are available for different functions and formatting, [see here for HTML options](http://rmarkdown.rstudio.com/html_document_format.html) and [here for PDF options](http://rmarkdown.rstudio.com/pdf_document_format.html). Rules in the header section will alter the whole document. Have a flick through quickly to familiarise yourself with the sorts of things you can alter by adding an option to the YAML header. 

Insert something like this at the top of your new `.Rmd` script:


``` {r eval=FALSE, purl=TRUE}
---
title: "Seagrass Biodiversity"
author: Jane Doe
date: 2020-02-03
output: html_document
---
```

By default, the `title`, `author`, `date` and `output` format are printed at the top of your HTML document. This is the minimum you should put in your header section. 

Now that we have our first piece of content, we can test the `.Rmd` file by compiling it to HTML. To compile your `.Rmd` file into a HTML document, click the **Knit** button in the taskbar.

By default, RStudio opens a separate preview window to display the output of your .Rmd file. If you want the output to be displayed in the Viewer window in RStudio (the same window where you would see plotted figures / packages / file paths), select "View in Pane" from the drop down menu that appears when you click on the **Knit** button in the taskbar, or in the _Settings gear icon_ drop down menu next to the **Knit** button. 

A preview appears, and a HTML file is also saved to the same folder where you saved your `.Rmd` file.

#### Code Chunks

Below the YAML header is the space where you will write your code, accompanying explanation and any outputs. Code that is included in your `.Rmd` document should be enclosed by three backwards apostrophes ```` ``` ```` (grave accents). These are known as code chunks and look like this:

```{r eval=FALSE, purl = TRUE}
norm <- rnorm(100, mean = 0, sd = 1)
```


Inside the curly brackets is a space where you can assign rules for that code chunk. The code chunk above says that the code is R code. We'll get onto some other curly brace rules later.

You can run an individual chunk of code at any time by placing your cursor inside the code chunk and selecting _Run -> Run Current Chunk_:

It's important to remember when you are creating an R Markdown file that if you want to run code that refers to an object, for example:

```{r eval=FALSE, purl = TRUE}
plot(dataframe)
```

you must include instructions showing what `dataframe` is, just like in a normal R script. For example: 

```{r eval=FALSE, purl = TRUE}
A <- c("a", "a", "b", "b")
B <- c(5, 10, 15, 20)
dataframe <- data.frame(A, B)
plot(dataframe)
```

Or if you are loading a dataframe from a CSV file, you must include the code in the `.Rmd`: 

```{r eval=FALSE, purl = TRUE}
dataframe <- read_csv(here("read_data", "mydata.csv")
```

Similarly, if you are using any packages in your analysis, you will have to load them in the `.Rmd` file using `library()` as in a normal R script. 

```{r eval=FALSE, purl = TRUE}
library(dplyr)
```

### Hiding code chunks

If you don't want the code of a particular code chunk to appear in the final document, but still want to show the output (e.g. a plot), then you can include `echo = FALSE` in the code chunk instructions. 


```{r, echo = FALSE}
A <- c("a", "a", "b", "b")
B <- c(5, 10, 15, 20)
dataframe <- data.frame(A, B)
plot(dataframe)
```

Similarly, you might want to create an object, but not include both the code and the output in the final `.html` file. To do this you can use, `include = FALSE`. Be aware though, when making reproducible research it's often not a good idea to completely hide some part of your analysis:


```{r, include = FALSE}
richness <- 
  edidiv %>%
    group_by(taxonGroup) %>%
    summarise(Species_richness = n_distinct(taxonName))
```


In some cases, when you load packages into RStudio, various warning messages such as "Warning: package 'dplyr' was built under R version 3.4.4" might appear. If you do not want these warning messages to appear, you can use `warning = FALSE`


```{r, warning = FALSE}
library(dplyr)
```


__REMEMBER: R Markdown doesn't pay attention to anything you have loaded in other R scripts, you MUST load all objects and packages in the R Markdown script.__

### Inserting Figures
Inserting a graph into `RMarkdown` is easy, the more energy-demanding aspect might be adjusting the formatting.

By default, `RMarkdown` will place graphs by maximising their height, while keeping them within the margins of the page and maintaining aspect ratio. If you have a particularly tall figure, this can mean a really huge graph. In the following example we modify the dimensions of the figure we created above. To manually set the figure dimensions, you can insert an instruction into the curly braces:


```{r, fig.width = 4, fig.height = 3}
A <- c("a", "a", "b", "b")
B <- c(5, 10, 15, 20)
dataframe <- data.frame(A, B)
plot(dataframe)
```


### Inserting Tables

#### Standard R Markdown

While R Markdown can print the contents of a data frame easily by enclosing the name of the data frame in a code chunk:


```{r}
dataframe
```


this can look a bit messy, especially with data frames with a lot of columns. Including a formal table requires more effort.

#### kable() function from knitr package

The most aesthetically pleasing and simple table formatting function I have found is `kable()` in the `knitr` package. The first argument tells kable to make a table out of the object `dataframe` and that numbers should have two significant figures. Remember to load the `knitr` package in your `.Rmd` file as well.


```{r}
library(knitr)
kable(dataframe, digits = 2)
```

## Data Quality Tests

Often, as we go through a data analysis we are checking the quality of the data as we go. You might, for example:
* Count the number of NAs in specific columns
* Check measurements for outliers
* Check sampling locations accuracy (are your sites on land?)
* Check all date and time formats are valid

It's often difficult to know at the end of an analysis project whether data quality has accidentally been compromised. It's also difficult to know what checks of the data have been done. Therefore it's a great idea to write tests into your data analysis as you go, such that if something becomes a problem, your diagnostic plots or tables will alert you!

```{r Data quality tests}
library(tidyverse)
library(lubridate)
library(obistools)

survey_data <- read_csv("https://raw.githubusercontent.com/HakaiInstitute/jsp-data/master/data/survey_data.csv")

seine_data <- read_csv("https://raw.githubusercontent.com/HakaiInstitute/jsp-data/master/data/seine_data.csv")

fish_field_data <- read_csv("https://raw.githubusercontent.com/HakaiInstitute/jsp-data/master/data/fish_field_data.csv")

fish_lab_data <- read_csv("https://raw.githubusercontent.com/HakaiInstitute/jsp-data/master/data/fish_lab_data.csv")

survey_seines_fish <- left_join(survey_data, seine_data) %>% 
  left_join(fish_field_data)
```

### Location check

```{r}
no_fish_loc <- survey_seines_fish %>% filter(
                                     is.na(survey_seines_fish$lat))
```

`r nrow(no_fish_loc)` fish have no location information associated with them.

Check if any of your points are on land.

```{r}
seine_data <- seine_data %>% 
  mutate(decimalLatitude = lat,
         decimalLongitude = long)

check_onland(seine_data)
land_points <- check_onland(seine_data)
#
plot_map(land_points, zoom = TRUE)
```


### Statistical QC

```{r}
# Lets do some stats
```

### Relational Joins

Do all seines have a survey to match to?

```{r}
#antijoin should yield zero
seines_match_surveys <-nrow(anti_join(seine_data, survey_data))

ifelse(seines_match_surveys == 0, "yes, seines match surveys", "error: no, some seines don't have matching surveys")
```

Do all lab fish have a matching field fish?

```{r}



lab_match_field <- nrow(anti_join(fish_lab_data, fish_field_data))

ifelse(lab_match_field == 0, "yes, all lab fish have a matching field fish", "error: no, some lab fish do not have a mathing field fish")
```

Do all fish have a matching seine?

```{r}
fish_data <- left_join(fish_field_data, fish_lab_data, by = "ufn")

n_umatched_fish <- nrow(anti_join(fish_data, seine_data, by = 'seine_id'))

ifelse(lab_match_field == 0, "yes, all fish have a matching seine they came from", "error: no, not all fish have a seine associated with it")
```

## Data Visualization Including Maps

First, load surveys_complete

```{r}
surveys_complete <- read_csv("https://raw.githubusercontent.com/Pelagic-Ecosystems/R-workshop/master/read_data/surveys_complete.csv")
```

### Plotting with **`ggplot2`**

**`ggplot2`** is a plotting package that makes it simple to create complex plots
from data in a data frame. It provides a more programmatic interface for
specifying what variables to plot, how they are displayed, and general visual properties. 

**`ggplot2`** functions like data in the 'long' format, i.e., a column for every variable,
and a row for every observation. Well-structured data will save you lots of time
when making figures with **`ggplot2`**

ggplot graphics are built step by step by adding new elements. Adding layers in
this fashion allows for extensive flexibility and customization of plots.

To build a ggplot, we will use the following basic template that can be used for different types of plots:

```
ggplot(data = <DATA>, mapping = aes(<MAPPINGS>)) +  <GEOM_FUNCTION>()
```

- use the `ggplot()` function and bind the plot to a specific data frame using the
      `data` argument

```{r, eval=FALSE, purl=FALSE}
ggplot(data = surveys_complete)
```

- define a mapping (using the aesthetic (`aes`) function), by selecting the variables to be plotted and specifying how to present them in the graph, e.g. as x/y positions or characteristics such as size, shape, color, etc.

```{r, eval=FALSE, purl=FALSE}
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length))
```

- add 'geoms' – graphical representations of the data in the plot (points, lines, bars). **`ggplot2`** offers many different geoms; we will use some 
  common ones today, including:
  
      * `geom_point()` for scatter plots, dot plots, etc.
      * `geom_boxplot()` for, well, boxplots!
      * `geom_line()` for trend lines, time series, etc.  


**Notes**

- Anything you put in the `ggplot()` function can be seen by any geom layers
  that you add (i.e., these are universal plot settings). This includes the x- and
  y-axis mapping you set up in `aes()`.
- You can also specify mappings for a given geom independently of the
  mappings defined globally in the `ggplot()` function.
- The `+` sign used to add new layers must be placed at the end of the line containing
the *previous* layer. If, instead, the `+` sign is added at the beginning of the line
containing the new layer, **`ggplot2`** will not add the new layer and will return an 
error message.

### **`ggplot2`** themes

In addition to `theme_bw()`, which changes the plot background to white, **`ggplot2`**
comes with several other themes which can be useful to quickly change the look
of your visualization. The complete list of themes is available
at <http://docs.ggplot2.org/current/ggtheme.html>. `theme_minimal()` and
`theme_light()` are popular, and `theme_void()` can be useful as a starting
point to create a new hand-crafted theme.

The
[ggthemes](https://jrnold.github.io/ggthemes/reference/index.html) package
provides a wide variety of options (including an Excel 2003 theme).
The [**`ggplot2`** extensions website](https://www.ggplot2-exts.org) provides a list
of packages that extend the capabilities of **`ggplot2`**, including additional
themes.

#### Faceting

* `facet_wrap()`
* `facet_grid()`

#### Arranging Multiple Plots

`cowplot()`

```{r}
library(tidyverse)

# The tidyverse library has a data frame object called mpg, it's about cars.
# Check it out
mpg

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

# What about this other column class? Maybe we want to see what type of car it is too

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))

# It would be nice to see a trend line

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))

# The default smoothing line is a loess model, which looks funny here, lets use a linear model

# It would be nice to see a trend line

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) +
  geom_smooth(mapping = aes(x = displ, y = hwy), method = lm)

ggsave(here("figs", "mpg.png"))
```


### Factors

Using `read_csv()` from `tidyverse` treats words as characters.

Using `read.csv()` from base treats words as factors; not great.

Long story short: Use read_csv and then you get to decide what's treated as a factor. 

[Be the boss of your factors](https://stat545.com/factors-boss.html)

```{r}
install.packages('gaminder')
```


### Dates and Times

`read_csv()` from `tidyverse` reads ISO standard date formats (yyyy-mm-dd) as date classes.

`read.csv()` from base R reads them as characters.

Another long story short: Use `read_csv()`

```{r}
class(fish_d09$date)
```

[Stat 545](https://stat545.com/date-time.html)


### Site Maps

Section Contributors:

* Dr. Daniel Okamoto
* Jenn Burt 

A common task amongst most field researchers is the need to make a basic site map to describe the location of your sampling. Using Arc GIS is the most common way to produce maps at Hakai, but sometimes a simple solution that could be implemented in R is desired. 

For a high resolution map with the resolution needed to see detailed coastline features you can use the following code and shapefile. To get this to work on your computer, download the shape file and put it in a R Studio project sub-folder called data.


```{r, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
###################################################### #
###  Script to make a BC map                       ### #
###  Author:  D.K. Okamoto (modified by Jenn Burt) ### #
###################################################### #

# Install the libs you need
# Libraries needed to run this code
library(raster)
library(maps) 
library(mapdata)
library(maptools)
library(mapproj)
library(rgeos)
library(rgdal)
library(ggplot2)
library(ggsn)
library(tidyverse)
library(here)
```

#### High Resolution Maps

The high resolution map used here requires that you download a set of ESRI shape files from this book's GitHUB repository. Those files can be downloaded from the `2_Shapefile` folder [here](https://github.com/HakaiInstitute/hakai_guide_to_r/tree/master/data). Put the `2_Shapefile` folder into the data folder of your R-Studio project.

This script assumes you are using the `here()` package in conjunction with R-Studio projects to obviate setting your working directory. 

```{r}
############## Make a map with sites ################## 
########## Using high resolution shapefile ################

BC.shp <- readOGR(here("read_data","2_Shapefile", "COAST_TEST2.shp"))

### chose the lat/long extent you want to show
Ncalvert <- extent(-128.18, -127.94, 51.61, 51.78)

### crop your shapefile polygons to the extent defined
# takes a moment to run (patience grasshopper)
BC.shp2 <- crop(BC.shp,Ncalvert)

### project and fortify (i.e. turn into a dataframe)
BC.df <- fortify(BC.shp2)

# (IF DESIRED) Load .csv file with your specific study site lat/longs
# this file is a dataframe with 4 columns: site_name, otterOcc(Y or N), lat, long  
# EXPT
# sites <- read.csv("/Users/jennb/Dropbox/Simple_BC_map/EXPTsites.csv", header = T)

# Jenn graph 
# here is where you can see the styles of north arrow (scroll to
# bottom): http://oswaldosantos.github.io/ggsn/ 
# the high resolution shape file works well at this scale 
# as it gives lots of the coastline detail
ggplot() + theme_bw() +
  geom_polygon(
    data = BC.df,
    aes(x = long, y = lat, group = group),
    colour = "black",
    size = 0.1,
    fill = 'grey95'
  ) +
  coord_cartesian(xlim = c(-128.17, -127.95),
                  ylim = c(51.63, 51.772)) +
  scalebar(
    BC.df,
    dist = 3,
    dist_unit = "km",
    st.size = 4,
    height = 0.01,
    transform = TRUE,
    model = 'WGS84',
    anchor = c(x = -127.96, y = 51.63)
  ) +
  north(
    data = BC.df,
    scale = 0.1,
    symbol = 3,
    anchor = c(x = -128.15, y = 51.775)
  ) +
  theme(
    panel.grid.minor = element_line(colour = NA),
    panel.grid.major = element_line(colour = NA),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10)
  )
```

```{r}
### if you want to make a larger Central coast map, just change the extent selected
CCoast <- extent(-128.48, -127.9, 51.5, 52.1)
# crop the map to the new extent
CC.shp2 <- crop(BC.shp,CCoast)
# fortify
CC.df <- fortify(CC.shp2)

# Jenn graph
fig1 <- ggplot() + theme_bw() +
  geom_polygon(
    data = CC.df,
    aes(x = long, y = lat, group = group),
    colour = "black",
    size = 0.1,
    fill = 'grey85'
  ) +
  coord_cartesian(xlim = c(-128.5, -127.95),
                  ylim = c(51.63, 52.05)) +
  scale_x_continuous(breaks = c(-128.4, -128.2, -128.0)) +
  scalebar(
    CC.df,
    dist = 5,
    dist_unit = "km",
    st.size = 3.5,
    height = 0.014,
    transform = TRUE,
    model = 'WGS84',
    anchor = c(x = -128.33, y = 51.64)
  ) +
  north(
    data = CC.df,
    scale = 0.07,
    symbol = 3,
    anchor = c(x = -128.465, y = 52.056)
  ) +
  theme(
    panel.grid.minor = element_line(colour = NA),
    panel.grid.major = element_line(colour = NA),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    legend.position = "none"
  )
fig1

# I use this code to export a nice PDF file of specific dimensions.
cairo_pdf("Fig1.pdf", width=4, height=5)
print(fig1)
dev.off()
```


#### Medium Resolution PBS Mapping Package

The Pacific Biological Station in Nanaimo has put together a mapping package that contains some medium resolution files of the Pacific Coast.

```{r}
############## Make a map with the sites ################## #
############# Using DFO coastline data file ################

#this is lower resolution than the shapefile above

library(PBSmapping)

## Plot the map
data(nepacLLhigh)       # DFO BC Coastline data - high resolution
plotMap(
  nepacLLhigh,
  xlim = c(-128.52, -127.93),
  ylim = c(51.56, 52.07),
  col = "grey90",
  bg = "white",
  tckMinor = 0,
  xlab = "",
  ylab = "",
  lwd = 0.5
)
box()

# add a scale bar
map.scale(
  x = -128.455,
  y = 51.61,
  ratio = FALSE,
  relwidth = 0.2
)
```

#### Low Resolution Pacific Coast Maps

For lower resolution maps to represent larger scales you can use the maps from the `maps` and `mapdata` packages.

```{r}
############## Pacific Coast Map ##################
################################################### #

# create a data file to make a basemap
# this database has a lower resolution (which is fine for large scale map)
m <- map_data("world", c("usa", "Canada"))

# this database has a way higher resolution
d <- map_data("worldHires", c("Canada", "usa", "Mexico"))

# make a basic map, all one colour
# play around with xlim and ylim to change the extent
ggplot() + geom_polygon(data = d, aes(x=long, y = lat, group = group)) + theme_bw()+
  coord_map("conic", lat0 = 18, xlim=c(210, 237), ylim=c(46,62))
```

```{r}
ggplot() +
  geom_polygon(
    data = subset(m, region == "Canada"),
    aes(x = long, y = lat, group = group),
    fill = "grey65",
    colour = "black",
    size = .1
  ) + theme_bw() +
  geom_polygon(
    data = subset(m, region == "USA"),
    aes(x = long, y = lat, group = group),
    fill = "white",
    colour = "black",
    size = .1
  ) +
  coord_map(
    "conic",
    lat0 = 18,
    xlim = c(195, 238),
    ylim = c(30, 62.5)
  ) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(colour = "grey"),
    #change "grey" to NA to remove
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )
```

```{r}
# playing with extent and colour
ggplot() +
  geom_polygon(data = subset(d, region == "Canada"), aes(x = long, y = lat, group = group)) +
  geom_polygon(data = subset(d, region == "USA"),
               aes(x = long, y = lat, group = group),
               fill = "darkgrey") +
  coord_map(xlim = c(195, 240), ylim = c(45, 65)) +
  theme_bw()
```

```{r}
# here you can see that if you use the other dataframe "d" the resolution is much higher. 
# this is good for smaller chunks of the BC coast, but less good for a PNW map
ggplot() +
  geom_polygon(
    data = subset(d, region == "Canada"),
    aes(x = long, y = lat, group = group),
    fill = "darkgrey",
    colour = "black",
    size = .1
  ) + theme_bw() +
  geom_polygon(
    data = subset(d, region == "USA"),
    aes(x = long, y = lat, group = group),
    fill = "white",
    colour = "black",
    size = .1
  ) +
  coord_map(
    "conic",
    lat0 = 18,
    xlim = c(195, 240),
    ylim = c(45, 61)
  ) +
  theme(
    panel.grid.minor = element_line(colour = NA),
    panel.grid.major = element_line(colour = NA),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )
```
