---
title: "R Markdown intro"
author: "Brett"
date: "04/02/2020"
output:
 html_document:
   theme: cosmo
   code_folding: show
   toc: true
   toc_float: true
   number_sections: true
---

YAML header ^ 

# YAML header options


```{r setup, include=FALSE}
# This is a code chunk comment using a pound sign
knitr::opts_chunk$set(echo = TRUE) # This makes all the code chunks set to echo = true
library(tidyverse)
```

Anything that's written here will render as text

# First Level heading
## Second level heading
### Third

* Bullet 1
* Bullet 2
* Bullet 3

_italics_
__bold__

R Markdown relies of Pandoc and LaTex under the hood

To insert a new code chunk use a shortcut:
* Mac: cmd-option-i
* WIndows: ctrl-alt-i

```{r code chunk intro}
5 + 5


# o comment multiple lines at once press cmd-shift-c
# dgf
# fdgsdfg
# sfgsdfgf
# fgsgfg

# To span multiple lines in R use ctrl-opt-up.
# Hit escape 

```

### Code chunks

Useful for controlling the output and evaluation of each code chunk. Useful options are eval, echo, and include.

#### Figures and tables in chunks

```{r}
b <- c(1,2,3,4)
a <- c("a", "a", "b", "b")

#use a new function called tibble to create a dataframe(aka tibble)
df <- tibble(a,b)

ggplot(df, aes(a,b)) +
  geom_point()

# create a nice looking table
knitr::kable(df, format = "markdown")

# use the alignment argument to adjust alignment options
```

# Data Quality Tests

## Check lat and lons

```{r message=FALSE, warning=FALSE}
# read in our salmon survey data from github
seine_data <- read_csv("https://raw.githubusercontent.com/HakaiInstitute/jsp-data/master/data/seine_data.csv")

#write a test to check for NAs in lat/lon
# R evaluates TRUE as numeric 1, and FALSE as 0
 
n_nas <- sum(is.na(seine_data$lat))
#For inline r object display always use `r object_name`
```

The number of lat values that are NA is: `r n_nas `

```{r relational joins}
survey_data <- read_csv("https://raw.githubusercontent.com/HakaiInstitute/jsp-data/master/data/survey_data.csv")

# lets make sure that all seines have a match for survey metadata

# different types of joins include full_join, left_join, but also anti_join.

# check which surveys don't have a corresponding seine
survey_seine_match <- anti_join(survey_data, seine_data, by = 'survey_id')

# check which seines don't have corresponding metadata from surveys
n_sad_seines <- nrow(anti_join(seine_data, survey_data))
```

The number of seines that don't have matching survey metadata is: `r n_sad_seines`

# Dates and times

This material comes from Jenny Bryans Stat 545 course materials at stat545.com


```{r}

library(lubridate) # library(tidyverse) doesn't load lubridate
Sys.Date()
today()

# check object class using base r to see if an object is stored as a date
str(Sys.Date())

Sys.time()
now()

str(Sys.time())
# POSIXct lets you know you're dealing with a date/time format

survey_data$survey_date
# Best practice is to store dates as YYYY-MM-DD

# Create month column from our survey_date column

survey_data <- survey_data %>% 
  mutate(month = month(survey_date),
         year = year(survey_date),
         yday = yday(survey_date),
         day = day(survey_date)) %>% 
  mutate(weird_date = paste(day, month, year, sep = "-")) %>% 
  mutate(ymdate = dmy(weird_date))

survey_data <- 
  
d1 <- "January 1, 2010"
mdy(d1)

d2 <- "2015-Mar-07"
ymd(d2)

d3 <- ""
```

### Time spans 

Type of spans of time:
* Duration: represents an exact number of seconds
* periods: represent human units like weeks, days or months
* intervals: represents a starting point and an end-point

```{r}
# How old is Brett
b_age <- today() - ymd(19860918)
class(b_age)

b_age <- as.duration(b_age)
b_age
class(b_age)

dseconds(15)
dminutes(10)

# Durations always record times in seconds


tomorrow <- today() + ddays(1)
tomorrow

last_year <- today() - dyears(1)

last_year
```

```{r}
# because durations always use seconds you get this:

one_pm <- ymd_hms("2016-03-12 13:00:00", tz = "America/New_York")

one_pm

one_pm + ddays(1)
# adding durations can sometimes cause problems as demonstrated

```

## Periods

Periods are time spans but don't have a fixed length in seconds. Instead they work like human times like days and months

```{r}
one_pm

one_pm + days(1) # ddays is a duration, days is a period class

```

Like durations, periods can be created with anumber of friendly constructor fucntions:

```{r}
seconds(15) # dseconds provides time span class of duration, whereas seconds provides time span class of period

minutes(10)
# you can multiply and add periods

# Leap year example:
ymd("2016-01-01") + dyears(1)
ymd("2016-01-01") + years(1)

```

### Intervals

```{r}
dyears(1) / ddays(365) # durations

years(1) / days(1) # periods

next_year <- today() + years(1)
next_year

# If you'd like to know the exact duraration in days between two times use an interval and divide by the duration you'd like (ddays)
# %--% is for showing a duration
(today() %--% next_year) / ddays(1)

# To find out how many periods fall into an interval, we need to use integer division.
(today() %--% next_year) %/% days(1)

```

