---
title: "Hakai Institute Bottle File Data Quality Report"
author: "Brett Johnson"
date: '`r date()`'
output:
 html_document:
   theme: cosmo
   code_folding: hide
   toc: true
   toc_float: true
   number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = FALSE)
library(tidyverse)
library(devtools)
library(lubridate)
library(knitr)
library(here)
library(car)
library(testthat)
library(fuzzyjoin)

# Read data (see data import script for data access code)
chl_all_cols <- read_csv(here("raw_data", "cha_all_cols.csv"), guess_max = 20000) 
nuts_all_cols <- read_csv(here("raw_data", "nuts_all_cols.csv"),
                          guess_max = 100000, 
                          col_types = cols(no2_no3_ugl = col_double(),
                                           tp = col_double())) 
poms_all_cols <- read_csv(here("raw_data", "poms_all_cols.csv"), guess_max = 20000) 


qu39_ctd <- read_csv(here("raw_data", "qu39_ctd.csv"), guess_max = 100000)

# Carbon file provided  in a file by Carrie. Should be replaced with citable version
carbon_chem <- read_csv(here("raw_data", "oa-random-file-from-Carrie.csv")) %>% 
  mutate(cc_date = ymd_hms(paste(DATE_UTC, TIME_UTC)))


# Define common columns I want back from every discrete sample
common_columns <- c("event_pk", "date",  "collected", "hakai_id",
                    "site_id", "line_out_depth", "sampling_bout",
                    "pressure_transducer_depth") # doesn't currently use lat and lon

site <- "QU39"

#6 Deal with bottom issues: sometimes CTD doesn't go all the way down, and if we use target depth of discrete sample to match, there will be no match.

#7 Replace CTD surface measurements with YSI measurements

#8 Add DOC and CO2 to bottle file

#TODO: Pair closest CTD depth to bottle depths.
  # [ ] Preferentially use solo depths rather than target depths
  # [ ] If no solo depths, use target depth
#TODO: include flag columns for each variable
#TODO: Look out for cases when there is no start_dt
#TODO: break-out files by event_pk just before putting into erddap
#TODO: eventually add sensible comment columns
#TODO: Confirm missingness with OSV?


#Note: xml has to be updated with new variable names each time another variable is added
```

# Discrete Samples

The point of this script is to join discrete sample results taken from Niskn bottle with phsyical and biogeochemical parameters from paired CTD casts. 

Along the way I write various quality checks to summarize the quality of various samples.

## Chlorophyll a


```{r Chl a}
chl <- chl_all_cols %>% 
  select(common_columns, filter_type, 
         chla_final, phaeo_final, chl_phaeo_quality_level = quality_level) %>% 
  filter(!filter_type %in% c('2 + 3', '20')) %>% 
  drop_na(filter_type, chla_final, phaeo_final) %>% 
  ungroup() %>% 
  filter(site_id == site)

collapse_all_rows <- function(df) {
  return(coalesce(!!! as.list(df)))
}

chl_wide <- chl %>% 
  pivot_wider(names_from = c(filter_type), values_from = c(chla_final, phaeo_final)) %>% 
  group_by(event_pk, line_out_depth) %>% 
  summarise_all(collapse_all_rows) %>% 
  select(event_pk:chl_phaeo_quality_level, "chla_final_Bulk GF/F", "chla_final_GF/F") %>% 
  ungroup()

```


```{r Nutrients}
nuts <- nuts_all_cols %>% 
  filter(site_id == site) %>% 
    select(common_columns, nh4_, no2_no3_um, no2_no3_ugl,
           no2_no3_units, tp, tdp, tn, tdn, srp, po4, sio2, po4pfilt, no3nfilt, po4punfl,
           no3nunfl, nh4nunfl, quality_level)
  
mean_nuts <- nuts %>% 
  group_by(event_pk, line_out_depth, collected) %>% 
  summarise_at(c("nh4_", "no2_no3_um", "no2_no3_ugl",
           "tp", "tdp", "tn", "tdn", "srp", "po4", "sio2", "po4pfilt", "no3nfilt", "po4punfl",
           "no3nunfl", "nh4nunfl"), funs(mean(., na.rm = TRUE), sd(., na.rm = TRUE), n())) # Ignores suspect replicates

```

## Particulate Organic Matter

```{r POMS}
poms <- poms_all_cols %>% 
  filter(site_id == site) %>% 
  select(common_columns, acidified, acidification_method, volume, part, screen_size, collected, preserved, analyzed, filter_portion, pre_weight_mg:row_flag, quality_level_poms = quality_level) %>%
  ungroup()

```

# Join Discrete Samples

## Join Chla with Nuts

```{r join discrete samples}

# This function assigns a new column thats a numeric of its date column. 
# This number is used to calculate euclidean distance across other numbers (including depth, and lat and long)
carbon_chem$numeric_collected <- as.numeric(carbon_chem$cc_date)
chl_wide$numeric_collected <- as.numeric(chl_wide$collected)
poms$numeric_collected <- as.numeric(poms$collected)
nuts$numeric_collected <- as.numeric(nuts$collected)
# create new depth column that gives preference to transducer depth

#preference to transducer depth
prf_2_td <- function(df) {
  df$priority_depth <- round(coalesce(df$pressure_transducer_depth, df$line_out_depth), 0)
  df
}

chl_wide <- prf_2_td(chl_wide)
poms <- prf_2_td(poms)
nuts <- prf_2_td(nuts)

chl_nuts <- distance_full_join(chl_wide, nuts, 
                                       by = c("priority_depth", "numeric_collected"), 
                               distance_col = "diff", method = "euclidean", max_dist = 1)

test_that("depths are the same", {
  chl_nuts <- drop_na(chl_nuts, c("priority_depth.x", "priority_depth.y"))
  expect_equal(chl_nuts$priority_depth.x, chl_nuts$priority_depth.y)
})

# fill in missing chl collection and depths with nuts depths and collection  times so I can joim other sample types all to chlorophyll
chl_nuts$numeric_collected.x <- coalesce(chl_nuts$numeric_collected.x,
                                         chl_nuts$numeric_collected.y)
chl_nuts$priority_depth.x <- coalesce(chl_nuts$priority_depth.x, 
                                      chl_nuts$priority_depth.y)

chl_nuts <- chl_nuts %>%
  rename_at(.vars  = vars(ends_with(".x")),
            .funs = funs(sub("[.]x$", "_chl", .))) %>%
  rename_at(.vars  = vars(ends_with(".y")),
            .funs = funs(sub("[.]y$", "_nuts", .)))
```

## Join Chl/Nuts with POMS

```{r}
#TODO: The problem is here, b/c some chl a rows will have no collection time or depth. So I need to coalesce the chl collected numeric with the nuts collected, and same for priority depth.
# The same problem with occur when I join carbon chem to chl_nuts_poms.
chl_nuts_poms <- distance_full_join(chl_nuts, poms,
                                    by = c("priority_depth_chl" =  "priority_depth",
                                           "numeric_collected_chl" =  "numeric_collected"), 
                               distance_col = "chl_poms_diff", method = "euclidean", max_dist = 1)

# rename common cols with poms suffix
chl_nuts_poms <- chl_nuts_poms %>%
  rename(event_pk_poms = event_pk,
         line_out_depth_poms = line_out_depth,
         date_poms = date,
         collected_poms = collected,
         hakai_id_poms = hakai_id,
         site_id_poms = site_id,
         sampling_bout_poms = sampling_bout,
         pressure_transducer_depth_poms = pressure_transducer_depth,
         numeric_collected_poms = numeric_collected,
         priority_depth_poms = priority_depth)

# fill in chla times and depths with POMS
chl_nuts_poms$numeric_collected_chl <- coalesce(chl_nuts_poms$numeric_collected_chl,
                                         chl_nuts_poms$numeric_collected_poms)

chl_nuts_poms$priority_depth_chl <- coalesce(chl_nuts_poms$priority_depth_chl, 
                                      chl_nuts_poms$priority_depth_poms)
```

### Join Chl_nuts_poms with carbon chem

```{r}
chl_nuts_poms_carbon <- distance_full_join(chl_nuts_poms, carbon_chem,
                                           by = c("priority_depth_chl" = "DEPTH_SAMPLING",
                                           "numeric_collected_chl" = "numeric_collected"))

chl_nuts_poms_carbon$priority_depth_chl <- coalesce(chl_nuts_poms_carbon$priority_depth_chl,
                                      chl_nuts_poms_carbon$DEPTH_SAMPLING)

#TODO: Rename carbon chem columns to match other discrete samples, including suffix on common columns

discrete_samples <- chl_nuts_poms_carbon

write_csv(discrete_samples, here::here("processed_data", "discrete_samples.csv"))
```

# Join Discrete Samples and CTD Depths

```{r BFG}
qu39_ctd$numeric_collected <- as.numeric(qu39_ctd$start_dt)

bottle_file <- distance_left_join(discrete_samples, qu39_ctd, by = c("priority_depth_chl" = "depth", "numeric_collected_chl" = "numeric_collected"))
#bottle_file <- left_join(discrete_samples, qu39_ctd, by = c("date", "hakai_id", "work_area", "year", "site_id" = "station", "line_out_depth" = "depth"))
```


```{r bf qc}
#create minimal bottle file to QC

hakai_ids <- bottle_file %>% 
  select(hakai_id, bottom_dt, date, date_chl, date_nuts)

depth <- bottle_file %>%
  as_tibble() %>% 
  select("pressure", "start_depth", "bottom_depth", "line_out_depth", "original_bottom_depth",
         "target_depth", "pressure_transducer_depth", "pressure_transducer_depth_chl",
         "pressure_transducer_depth_nuts")

dates_qc <- bottle_file %>% 
  as_tibble() %>% 
  select(ctd_file_pk, ctd_cast_pk, ctd_data_pk, bottom_dt, original_bottom_dt, start_dt, original_start_dt, year, date, event_pk, date_chl, collected_chl, date_nuts, collected_nuts, collected_poms)  # No date_poms?
  
test_that("years are the same", {
  expect_equivalent(year(dates_qc$original_bottom_dt), year(dates_qc$date_chl))
  expect_equivalent(day(dates_qc$original_bottom_dt), day(dates_qc$date_chl))
})  


#TODO: Figure out why there's no date_poms  
#TODO: Convert all the dates to durations

min_bf <- bottle_file %>% 
  select(event_pk, line_out_depth, date, end_dt, temperature, collected_chl, collected_nuts, collected_poms)

ctd_time_match_qc <- min_bf %>% 
  mutate(median_discrete = median(c(collected_nuts, collected_chl)),
         ctd_time_diff = abs(difftime(end_dt, median_discrete, units = "mins")),
         ctd_time_flag = ifelse(ctd_time_diff > 180, "SVC", "")) %>% 
  filter(ctd_time_flag != "")
  
  

kable(ctd_time_match_qc, caption = "Table. Events for which CTD collection time was more than 180 minutes")

```
